\cleardoublepage
\chapter*{Acknowledgements}

\thispagestyle{NoHeader}

First and foremost, I wish to express my greatest gratitude to Prof. Dr. Cornelis (Kees) Dullemond,
for acting as my supervisor while I was writing this thesis, and for providing me with invaluable 
guidance, advice and feedback. \\

Second, I want to thank Prof. Dr. Carsten Dominik, who kindly took on the role of second 
examiner. \\

Also, since this thesis probably marks the end of my time as a student in Heidelberg, it needs to 
be said that I am extremely grateful for the last few years that I could spend in this beautiful 
city, and for all the learning opportunities granted to me by \textit{Ruperto Carola} University of 
Heidelberg, the Max Planck Institute for Astronomy, and especially the planet formation group and 
its members at the Heidelberg Institute for Theoretical Astrophysics. \\

On a more personal note, I wish to thank my mother, father, and grandmother, who kindled my 
curiosity and interest into the natural sciences from an early age on. Also, in that context I 
want to mention Kati, Jürgen, and Alexander, as well as my friends and family at DPSG
Ulm-Söflingen. \\

Without the enduring support of the mentioned individuals over many years, this thesis might have 
never come to fruition. As such, it is dedicated to them.

% This thesis is dedicated to them.

% Mother, father, grandmother
% Kati, Juergen, Alexander
% friends at DPSG Ulm-Söflingen

% family and friends 


% friends at DPSG Ulm-Söflingen for their support over the years.

% I wish to thank my mother for making it possible to study physics in Heidelberg,
% as well as my father and grandmother, who from an early age on inspired me to be curious about 
% the Universe. \\

% Specifically 
% Carolin (Lina) Kimmig,
% who was kind enough to look over my thesis before handing it in, and providing me with feedback. 
% \\

% Third, it needs to be said that I am incredibly grateful to the Ruperto Carola University of 
% Heidelberg. \\


% Lastly, I'd like to give a shoutout to my friends at DPSG Ulm-Söflingen.

% Of course 
% Lastly, I want to thank 
% Lastly, I want to give a quick shoutout to my friends at DPSG Ulm-Söflingen.
% Second, I want to thank the planet formation group and its members, ... \\
% Third, it needs to be said that I am very grateful to 
% have been a part of the planet formation group, as well as to its members,

% \begin{itemize}
    % \item Prof. Dr. Cornelis (Kees) Dullemond
    %     - for being my supervisor
    % \item Heidelberg University
    % \item Heidelberg Institute for Theoretical Astrophysics
    % \item Prof. Dr. Ralf Klessen
    %     - for inspiring me to study computational physics in his lecture
    % \item my parents 
        % Angelika Schäuffelen and Christian Mader (and grandmother)
    % \item family and friends
    % \item shoutout to DPSG Ulm-Söflingen
% \end{itemize}
