Our first task will be the definition of a model for the proto-planetary disk, as well as for
the properties of its gas and dust contents. It needs to be said here that due to the complexity of
the involved processes and the (literally) astronomically large number of particles present in the 
disk, our model can only ever be a highly simplified approximation to reality. \\

Nonetheless, the goal of the following chapter(s) (?) will be to formulate a model that captures a 
typical PPD's most relevant properties, which would then allow us to conduct studies of the 
processes involved in dust particle coagulation and/or fragmentation in the subsequent chapters. 
To be able to construct such a framework, it will be necessary to make a couple of assumptions:

% \todo{Do what? (Star, Disk, Gas, Dust, Coagulation/Fragmentation)} \\

% First Assumptions {{{
\section{First Assumptions}

    \begin{enumerate}
        \item \assume{} First of all, we will base our model on the aforementioned nebular 
            hypothesis: The origin of the modeled proto-planetary disk is assumed to be found in 
            the collapse of a giant interstellar cloud. % consisting of mostly gas and dust. 
            As such, the disk's elementary and material composition is assumed to equal that of the 
            interstellar medium as well. \\
            \todo{Elements?}
    
        \item \assume{} Also, we will assume that the cloud collapse happened sufficiently long ago 
            that a star could already form at the center of the disk (while not long enough for 
            planet formation to occur).
            % (\todo{How long ago?}) [cite subsection]
            For the studies that will be carried out in the context of this thesis, the star's most
            relevant properties will be its mass $M_*$ as well as its luminosity $L_*$. 
            \assume{}
            Both of these properties will be assumed to equal those of our own solar system's Sun. 
            Thus, we define
            \begin{align}
                M_*&:=M_\odot=\SI{1.989e30}{\kg},\ \text{and} \\
                L_*&:=L_\odot\ =\SI{3.828e26}{\joule\second^{-1}},
            \end{align}
            where $M_\odot$ and $L_\odot$ label the solar mass and luminosity, respectively.
            % \todo{What type of star is it? Where in the main sequence is it? How old is it?} 
            % \begin{itemize}
            %     \item G-type main sequence star
            %     \item often imprecisely called "yellow dwarf" 
            %           (imprecisely, because its light is actually white)
            % \end{itemize}
            % \todo{The Sun was not as it is today when it formed $\approx$ 4.6 billion years ago.} \\
    
        \item
            We will assume most of the original gas cloud's mass to be concentrated inside this 
            star, while the disk surrounding it holds only a much smaller fraction of the total 
            mass. We label the stellar mass $M_*$, and the disk mass $M_\text{disk}$. 
            The ratio between the two shall be defined as
            \begin{equation}
                q_\text{disk-to-star}
                    =\frac{M_\text{disk}}{M_*}  
                    :=0.01
            \end{equation}
    
        \item \assume{Furthermore, the disk's mass is assumed to be dominated by the contribution 
            of circumstellar gas, with only a comparatively tiny contribution of dust particles.}
            Let $M_\text{dust}$ and $M_\text{gas}$ label the total mass present in the form of dust 
            and gas, respectively. 
            With
            \begin{equation}
                M_\text{disk} = M_\text{gas} + M_\text{disk},
            \end{equation}
            we define the dust-to-gas ratio as
            \begin{equation}
                q_\text{dust-to-gas}=\frac{M_\text{dust}}{M_\text{gas}}:=0.01
            \end{equation}
    
        \item The next assumption we will make concerns itself with disk symmetry. For this, let us 
            adopt cylindrical coordinates $(r, \varphi, z)$. In reality, each of the disk's 
            position-dependent properties could best be described by a function $f(r, \varphi, z)$ 
            that depends on all three of these coordinates. \\
            To keep things simple though, we will not attempt to resolve the disk in all three 
            spatial dimensions. Instead, we will focus mostly on the radial dependence of the 
            disk's features.
    
            On large scales, we will assume the disk's features to not depend significantly on 
            $\varphi$ and $z$.
    
            \todo{Consider a disk with an inner and outer disk radius
                $r_\text{min}=\SI{0.01}{\astronomicalunit}$ 
            and 
                $r_\text{max}=\SI{100}{\astronomicalunit}$,
            respectively.} \\

            \todo{Assume perfect radial symmetry?}
    
        \item \assume{Gas mean particle mass.}
            \begin{equation}
                m_g
                    := 2.3 \cdot m_p
            \end{equation}
            Here, $m_p$ labels the mass of a single proton. \\
    
        \item \assume{Gas is in hydrostatic equilibrium.}
            \begin{equation}
                \deriv{P}{z} = - \rho_g \cdot g_z  % TODO Why `g_z` and not `g` ?
            \end{equation}
    
        \item \assume{Gas behaves isothermically.}
            \begin{equation}
                P = \rho_g \cdot c_s^2
            \end{equation}
    
    \end{enumerate}
    
    \todo{Plot top-down view of the disk.}
    
    \input{diagrams/disk_structure.tex}

% }}}
% Radial Profile of the Midplane Temperature {{{ 
\clearpage\section{Radial Profile of the Midplane Temperature}

    \assume{To keep things simple, let us make the assumption that the gas and dust particles 
    in the disk behave like perfect black-body objects.} As such, the relationship between the
    thermal radiation flux $B$ coming from the star (i.e. the total energy per unit surface area 
    per unit time) and the temperature $T$ is given by the Stefan-Boltzmann law, which states that
    \begin{equation}
        \label{eq:stefan_boltzmann_law}
        B(T) = \sigma_\text{SB}\cdot T^4   % TODO Is this true?
    \end{equation}

    Here, $\sigma_\text{SB}$ labels the Stefan-Boltzmann constant, which is given by 
    \begin{equation}
        \sigma_\text{SB}
            = \frac{2\pi k_B^4}{15h^3 c^2}
            \approx \SI{5.670e-8}{\ \watt\ \kelvin^{-1}\ \meter^{-2}}
    \end{equation}

    In general, the relationship between the luminosity $L$, which labels the total emitted energy 
    per time, and the radiation flux $B$ is given by the surface integral
    % The luminosity, i.e. the total energy emitted per time, can be determined by evaluating the
    % integral of the radiation flux over the entire surface of the considered absorver/emitter.
    % Thus, the relationship between radiation flux $B(T)$ and luminosity $L(T)$ is given by
    \begin{equation}
        L(T) = \int_A B(T) \ \text dA
    \end{equation}
    
    \assume{
    Since we are assuming spherical symmetry (isotropy) of the star's radiative power output, we 
    can simplify this expression. For a fixed stellar luminosity $L_*$, the radiative flux $B(r)$ 
    at a distance $r$ from the star can be written as}
    \begin{equation}
        L_* = B(r) \cdot 4\pi r^2
    \end{equation}

    Plugging in \cref{eq:stefan_boltzmann_law} and rearranging for the temperature $T$ leads to
    \begin{equation}
        T(r) = \left(
            \frac{L_*}{4\pi r^2 \cdot \sigma_\text{SB}}
        \right)^{1/4}
    \end{equation}

    \todo{Inclusion of $f/2$:}
    \begin{equation}
    \boxed{
        T(r) = \left(
            \frac{f}{2}\cdot
            \frac{L_*}{4\pi r^2 \cdot \sigma_\text{SB}}
        \right)^{1/4}
    }
    \end{equation}

    \input{diagrams/midplane_temperature_profile.tex}

    % \begin{itemize}
    %     \item midplane temperature $T_\text{mid}(r)$ (b.c. isothermal) \\
    %         \todo{Plot $T_\text{mid}$ vs. $r$ (lin-log?)}
    %     \item thermal velocity $u_\text{th}$ \\
    %         \todo{Plot $u_\text{th}$ vs. $r$ (lin-log?)}
    %     \item sound speed (here?)
    % \end{itemize}

% }}}
% Radial Profile of the Gas Density {{{ 
\newpage\section{Radial Profile of the Gas Density}

    Our next task will be to formulate a simple model for the spatial distribution of the disk's 
    gaseous contents. In this context, we will adopt cylindrical coordinates $(r, \varphi, z)$ and
    define both the surface and volume density profiles along the radial axis $r$.

    % Radial Profile of the Gas Surface Density {{{ 
    \subsection{Radial Profile of the Gas Surface Density}

        The gas surface density $\Sigma_g(r,\varphi)$ is directly related to the gas volume density
        $\rho_g(r,\varphi,z)$, and can be derived from it by integrating the volume density from 
        $-\infty$ to $+\infty$ along the $z$-axis, i.e. along the axis perpendicular to the disk:
        \begin{equation}
            \label{eq:relationship_between_gas_surface_density_and_gas_volume_density}
            \Sigma_g(r,\varphi)
                = \int\limits_{-\infty}^{\infty} \rho_g(r,\varphi,z)\ \dd z
        \end{equation}

        \assume{As noted before in (cite), we will assume the disk to be perfectly radially 
        symmetric.} (\todo{At least macroscopically!}) \\
        For the density distributions, this means that
        \begin{align}
            \Sigma_g(r, \varphi) 
                &= \Sigma_g(r) 
            \ \ \ \text{and}\ \ \
            \\
            \rho_g(r, \varphi) 
                &= \rho_g(r)
        \end{align}
        Thus, we can simplify 
        \cref{eq:relationship_between_gas_surface_density_and_gas_volume_density} and write:
        \begin{equation}
            \Sigma_g(r)
                = \int\limits_{-\infty}^{\infty} \rho_g(r,z)\ \dd z
        \end{equation}
        
        \todo{So far, we do not yet have an expression for the gas volume density $\rho_g$.}
        Therefore, we can not derive the surface density $\Sigma_g$ by evaluating the above
        integral. Instead, we will go the other way round by making an ansatz for the surface gas
        density and then derive the volume density from that. \\

        In order to do this, we will 
        follow the work of Brauer, Dullemond, \& Henning \cite{brauer_dullemond_henning_2007},
        % Following the works of \cite{brauer_dullemond_henning_2007}
        % , \cite{andrews_2007}, and \cite{kitamura_2002}, 
        and make the assumption that the radial profile of the gas surface density can be 
        approximated by using an inverse power law of the form
        \begin{equation}
            \label{eq:gas_surface_density_profile_inverse_power_law_ansatz}
            \Sigma_g(r)=\Sigma_0\cdot\frac{1}{r^\sigma}
            \ \ \ \text{with}\ \ \
            \sigma:=0.8
        \end{equation}
        % TODO Why 0.8 ?

        \todo{Before we can use this, we need to define $\Sigma_0$}. \\

        The total gas mass present in the disk is given by $M_\text{gas}$, which we defined in 
        (\todo{cite}). \\
        \todo{Since the integration of the surface density over the entire disk surface must yield 
        this value, we can formulate the following condition:}
        \begin{align}
            M_\text{gas}
                % &= \int_{A_\text{disk}} \Sigma_g(r) \ \dd A \\
                &= \int\limits_{r_\text{min}}^{r_\text{max}} \Sigma_g(r) \cdot 2\pi r \ \dd r
        \end{align}
        % TODO Define r_min and r_max before.

        \todo{Above (cite): Discretize $r$-axis into $\mathcal N_r$ "bins"}
        \begin{equation}
            M_\text{gas}
                = \sum_{i=1}^{\mathcal N_r} \Sigma_g(r_i) \cdot 2\pi r_i \cdot \Delta r_i
        \end{equation}
        % TODO Define N_r before.

        Now we can plug in \cref{eq:gas_surface_density_profile_inverse_power_law_ansatz} and 
        rearrange for $\Sigma_0$:
        \begin{equation}
            \Sigma_0
                = M_\text{gas} \cdot \left[
                    \sum_{i=1}^{\mathcal N_m} 2\pi r_i^{1-\sigma} \cdot \Delta r_i
                \right]^{-1}
        \end{equation}
        % TODO Do this with dA instead of 4pi r dr ?

    % }}}
    % Radial Profile of the Gas Volume Density {{{ 
    \newpage\subsection{Radial Profile of the Gas Volume Density}
        
        For the derivation of the volume density's radial profile, we will make

        \assume{Isothermal condition} [cite] \\
        \assume{Disk is in hydrostatic equilibrium [cite]} \\

        Let us adopt cylindrical coordinates and consider a point $\vec r=(r,\phi,z)^T$ in the 
        disk. \\  % TODO Write with transposed `T` ?

        \assume{The disk is "much more wide than tall": (it is a disk, afterall...)}
        \begin{equation*}
            r\gg z 
        \end{equation*}

        \todo{Sketch disk} \\
        \input{diagrams/force_balance_in_disk_with_hydrostatic_equilibrium.tex}

        The acceleration due to Newtonian gravity is given by
        \begin{equation}
            g=\frac{GM_*}{r^2+z^2}
        \end{equation}

        Making use of the laws of trigonometry, we can write
        \begin{equation}
            g_z=g\cdot\sin\theta
        \end{equation}
        and
        \begin{equation}
            \sin\theta=\frac{z}{\sqrt{r^2+z^2}}
        \end{equation}

        Plugging all of this into [cite: above], we can write
        \begin{align}
            \deriv{P}{z}
                &= -\rho_g \cdot g_z \\
                &= -\rho_g \cdot g \cdot \sin\theta \\
                &= -\rho_g \cdot \frac{GM_*}{r^2+z^2} \cdot\frac{z}{\sqrt{r^2+z^2}} \\
                &= -\rho_g \cdot z \cdot\frac{GM_*}{ (r^2+z^2)^{\frac{3}{2}} } \\
            \Rightarrow
            \deriv{P}{z}
                &\approx -\rho_g \cdot z \cdot \Omega_K^2
        \end{align}

        For the last step, we made use of the fact that $r \gg z$, which allows us to make the 
        approximation
        \begin{equation}
            \frac{GM_*}{(r^2+z^2)^\frac{3}{2}}
                \approx \frac{GM_*}{r^3}
                = \Omega_K^2
        \end{equation}

        \todo{Differentiate isothermal condition with respect to $z$:}
        \begin{equation}
            \deriv{P}{z}
                = \deriv{\rho_g}{z}\cdot c_s^2
        \end{equation}

        \todo{We now have two expressions for $\pderiv{P}{z}$}, in [cite eq 1] and [cite eq 2]. \\
        \todo{Equating both of these and solving for $\pderiv{\rho_g}{z}$ leads to the following 
        differential equation:}
        % TODO Is it an ODE or a PDE ? Should I write `dy/dx` here, or `del y / del x` ?
        \begin{equation}
            \deriv{\rho_g}{z}
                = - \left(\frac{\Omega_K}{c_s}\right)^2 \cdot \rho_g \cdot z
        \end{equation}

        This differential equation can be solved by making the ansatz of a Gaussian distribution
        \begin{equation}
            \rho_g(z)
                = 
                    \rho_0\cdot\exp\left(-\frac{z^2}{2H_p^2}\right)
        \end{equation}
        where
        \begin{equation}
            H_p
                :=\frac{\Omega_K}{c_s}
        \end{equation}
        is the so-called \textit{pressure scale height} of the disk. \\
        % TODO socalled vs so-called

        \todo{Now that we know the behavior of the gas volume density, we can easily derive it from 
        the gas surface density like this:} 
        \begin{equation}
            \rho_g(r)
                =\frac{\Sigma_g(r)}{\sqrt{2\pi}\cdot H_p}
        \end{equation}
        \todo{Why can we do this?}

    % }}}
    % Radial Profile of the Gas Number Density {{{ 
    \subsection{Radial Profile of the Gas Number Density}

        \todo{Derive $N_g(r_i)$} \\

        \begin{equation}
            N_g
                =\frac{\rho_g}{m_g}
        \end{equation}

        \todo{Is this really even needed? (Don't I only need dust number density?)} \\

        \todo{Plot gas number density $N_g(r_i)$ vs $r_i$.} \\  % `N_g -> N`

    % }}}
    % Plots of Density Profiles {{{

        \newpage
        \todo{Plot gas surface density $\Sigma_g(r)$ vs $r$.} \\
        \todo{Plot gas volume density $\rho_g(r_i)$ vs $r_i$.} \\
        \todo{Plot $\rho_g(z)$.}
        \input{diagrams/gas_density_profiles.tex}

    % }}}

% }}}
% Gas Particle Kinematics {{{ 
\newpage\section{Gas Particle Kinematics}

    \todo{There are both stochastic and systematic contributions to the gas kinematics.}

    % Thermal Gas Velocity {{{ 
    \subsection{Thermal Gas Velocity}

        \todo{Sound speed:}
        \begin{equation}
            c_s = \sqrt{\frac{k_BT}{m_g}}
        \end{equation}

        \todo{Thermal velocity:} % TODO thermal SPEED
        \begin{equation}
            u_\text{th} = \sqrt{\frac{8}{\pi}} \cdot c_s
        \end{equation}

        \todo{Mean free path:}
        \begin{equation}
            \lambda_\text{mfp} = \frac{1}{n \cdot \sigma_{\text{H}_2}}
        \end{equation}

        \input{diagrams/thermal_gas_velocity.tex}

    % }}}
    % Radial Gas Velocity {{{ 
    \newpage\subsection{Radial Gas Velocity}

        \todo{Plot radial profile of "speed" (thermal + ..?)}

        The viscous evolution of the gas disk can be described by the continuity equation 
        \cite{birnstiel_dullemond_brauer_2010}:
        \begin{equation}
            \pderiv{\Sigma_g}{t}-\frac{1}{r}\pderiv{}{r}\bigg(\Sigma_g \cdot r \cdot u_g\bigg) = S_g
        \end{equation}
        Here, $u_g$ labels the \textit{radial gas velocity}.
        \\
        % TODO: What is $S_g$
        % TODO: Does $u_g$ label radial velocity or speed?
        % TODO: Choose either $uor and $v$ as variable names for velocities.
        %       Be consistent across the entire thesis.

        A solution to this equation is given by \cite{lynden-bell_pringle_1974}, where the 
        radial velocity of the gas is given by
        \begin{equation}
            u_{g}
            =-\frac{3}{\Sigma_g\cdot\sqrt{r}}
                \cdot\pderiv{}{r}\bigg(\Sigma_g \cdot \nu_g \cdot \sqrt{r}\bigg)
        \end{equation}
        % TODO: Have I already defined the viscosity somewhere? Else define here.

        \todo{Viscosity:}
        \begin{equation}
            \nu_g = \frac{1}{2} \cdot u_\text{th} \cdot \lambda_\text{mfp}
        \end{equation}

        \input{diagrams/radial_gas_velocity.tex}

    % }}}
% }}}

% \section{Symmetry Properties of the Disk}  
%     \todo{Rename section? $\to$ "Disk Structure" ?} \\
%     \todo{Plot disk structure (flaring disk, flaring index)} \\
%     \todo{What exactly is on the ordinate? ($r$ on the abscissa)} \\
%     \assume{Disk is radially symmetric: $f(r,\varphi,z)=f(r,z)$ (cylindrical coordinates)} \\
